---
title: "Covid Model Analysis"
author: "Joseph Heffner"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
  word_document: default
  pdf_document: default
---

# Setup

To run this analysis script you'll have to have the following packages installed (make sure `tidyverse` is updated): 

`here`, `conflicted`, `tidyverse`, `knitr`, `broom.mixed`, `kableExtra`, `ggrepel`, `psych`, `caret`, `corrplot`, `tableone`, `sjPlot`, `papaja`, `corrr`, `ggraph`, `igraph`

To install these packages simply use the `install.packages()` function and put each package name in as a character. 

**Note**: If you've already used the `here()` function in another script, you will have to open a new instance of R to ensure the relative pathing works for this script (or manually change the pathing). 

# Covid Models

The order of this Markdown will follow the order of results in the manuscript which is published in the journal of Humanities & Social Sciences Communication at the following [DOI TO ADD ONCE PUBLISHED]().

```{r setup, include=FALSE}
# Knitr options
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

reg_ouput <- FALSE    # set to TRUE if want to knit document with regression table output
cor_output <- FALSE   # knitting will not work with these graphs     

# Libraries
library(here)         # relative path
library(conflicted)   # resolve conflicts
library(tidyverse)    # tidy functions
library(knitr)        # knit functions
library(broom.mixed)  # tidy()
library(kableExtra)   # extra markdown functions
library(ggrepel)      # geom_text_repel
library(psych)        # alpha(), omega()
library(caret)        # train-test functions
library(corrplot)     # correlation plot
library(tableone)     # tables
library(sjPlot)       # regression tables
library(papaja)       # apa regression tables, requires an install from github
library(corrr)        # correlations
library(ggraph)       # graph visualizations
library(igraph)       # graph visualizations


# Resolve conflicts
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
```

# Data

There are two data files: `covid_models_data` which holds all variables used in the cross-validated model of emotional distress and `covid_emo_distress` which holds the item-level responses for the emotional distress scale. See manuscript and supplement for more detailed information about all the questionnaires included. 

```{r read_data}
# Relative paths
dir_parent <- here()
dir_data <- str_c(dir_parent, "/data")
dir_graphs <- str_c(dir_parent, "/graphs")

# Data
df_models <- read_csv(str_c(dir_data, "/covid_models_data.csv")) %>% select(-sub) 
df_emo_distress <- read_csv(str_c(dir_data, "/covid_emo_distress.csv")) %>% select(-sub)
```

## Covid models data

Below you can see a snippet of the data where the columns refer to: (if you want subject numbers you can keep them in the previous code chunk but they need to be removed for the modeling). 

```{r}
df_models %>% head() %>% 
  kable() %>%
  kable_styling()
```

**Note** All variables have been scaled and have a subscript (`_s`) which indicates this. 

- `erq_cog_reap_s`: cognitive reappraisal subscale of the emotion regulation questionnaire 
- `erq_exp_supp_s`: expressive suppression subscale of the emotion regulation questionnaire 
- `income_s`: ordinal income (see Supplement for scale)
- `tas_total_s`: total score for the toronto-20 alexithymia scale
- `cesd_total_s` total score for the center for epidemiological studies depression scale
- `gad_total_s`: total score for generalized anxiety disorder (7-item) scale
- `irq_total_s`: interpersonal regulation questionnaire total scale
- `bfi_extraversion_s`: extraversion subscale of the big five inventory scale 
- `bif_neuroticism_s`: neuroticism subscale of the big five inventory scale
- `ius_total_s`: intolerance of uncertainty scale
- `liberalism_s`: political ideology identification item
- `age_s`: subject age
- `covid_knowledge_s`: score from quiz about covid knowledge at the time of data collection (see Supplement) 
- `covid_estimate_log_s`: log-scaled estimate of current covid cases in the US at the time of data collection 
- `covid_behave1_s`: covid behaviors scale 1 (see Supplement)
- `covid_behave2_s`: covid behaviors scale 2 (see Supplement)
- `covid_media_`: media consumption about covid on relevant platform (`_facebook`, `_youtube`, `_twitter`, `_other`, `_govt`, `_online`, `_tv`, `_usa` [usa specific news], `_int`, [international specific news]) 
- `gender`: 0 - male, 1 - female

## Covid emotion distress data

See Supplement for corresponding item-level questions. 

# Stepwise CV model 

## Train model

```{r fit_model}
#### Test-train split
set.seed(123)

index_train <- as.numeric(caret::createDataPartition(y = df_models$covid_emo_distress_s, p = 0.75, list = FALSE, times = 1))
df_train_raw <- df_models[index_train, ]
df_test_raw <- df_models[-index_train, ]

#### Stepwise lm 
df_train_lm <- df_train_raw
df_test_lm <- df_test_raw

numCores <- parallel::detectCores() - 1 # never want to use all cores 
cl <- parallel::makeCluster(numCores, type = "SOCK")
doSNOW::registerDoSNOW(cl) # register cluster so caret knows to train in parallel

train_control <- trainControl(method = "cv", number = 10)
lm_fit <- train(covid_emo_distress_s ~ ., 
                data = df_train_lm, 
                method = "lmStepAIC", 
                trControl = train_control, trace = FALSE, # prevent output
                direction = "both") # forward and backward
parallel::stopCluster(cl) # stop cluster after training
```

## Test model

```{r test_model}
#### Test model
df_test_lm$model_fits <- predict(lm_fit, newdata = df_test_lm)
df_test_results <- postResample(pred = df_test_lm$model_fits, obs = df_test_lm$covid_emo_distress_s)
cor_value <- cor(df_test_lm$model_fits, df_test_lm$covid_emo_distress_s)
var_value <- cor_value^2

#### Calibration check 
m1 <- lm(covid_emo_distress_s ~ model_fits, data = df_test_lm)
#summary(m1)

df_test_plot <- ggplot(df_test_lm, aes(x = model_fits, y = covid_emo_distress_s)) + 
  geom_point() + 
  geom_smooth(method = "lm", se = F) + 
  scale_y_continuous(name = "Real COVID-19\nEmotion Distress", limits = c(-2.2, 2.4)) + 
  scale_x_continuous(name = "Predicted COVID-19\nEmotion Distress", limits = c(-2.2, 2.4)) + 
  coord_fixed() + 
  geom_hline(aes(yintercept = 0), linetype = "dashed") + 
  geom_vline(aes(xintercept = 0), linetype = "dashed") + 
  annotate(geom="text", y = 0.2, x = 1.75, label = str_c("r=", round(cor_value, 2)), color = "blue", hjust = 0) + 
  annotate(geom="text", y = 0.4, x = 1.75, label = str_c("r2=", round(var_value, 2)), color = "red", hjust = 0) + 
  theme_classic()
df_test_plot
ggsave(filename = str_c(dir_graphs, "/fig2a.pdf"), plot = df_test_plot, width = 4, height = 4)

#### Refit on entire data set
best_model <- as.character(lm_fit$finalModel$call[2]) %>% str_replace(., ".outcome", "covid_emo_distress_s")
final_model <- lm(formula = as.formula(best_model), data = df_models)
final_summary <- summary(final_model)

final_results <- as.data.frame(final_summary$coefficients) %>%
  tibble::rownames_to_column(var = "coefficient") %>%
  mutate(p_values = case_when(`Pr(>|t|)` <= .001 ~ "***", 
                              `Pr(>|t|)` <= .01 ~ "**", 
                              `Pr(>|t|)` <= .05 ~ "*", 
                              `Pr(>|t|)` > .05 ~ "n.s.")) %>%
  filter(coefficient != "(Intercept)")

final_betas <- ggplot(final_results, aes(x = reorder(coefficient, -abs(Estimate)), y = Estimate)) + 
  geom_bar(stat = "identity", position = position_dodge(.9), fill = "white", color = "black") + 
  geom_errorbar(aes(ymin = Estimate - `Std. Error`, ymax = Estimate + `Std. Error`), width = 0) + 
  geom_text(aes(label = p_values, y = Estimate + .1)) +
  xlab("Coefficient Names") + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust= 1))
final_betas
ggsave(filename = str_c(dir_graphs, "/fig2b.pdf"), plot = final_betas, width = 7, height = 4)

#### Export regression results
if (reg_ouput == TRUE) {
  
  # sjPlot
  tab_model(final_model, 
            pred.labels = c("Intercept", "Alexithymia", "Anxiety", "Interpersonal (IRQ)", 
                            "Extraversion", "IUS", "Liberalism", "Age", "Covid Knowledge", 
                            "Covid Behavior", "Covid Media: Facebook", "Covid Media: Twitter", 
                            "Covid Media: TV", "Covid Media: USA", "Covid Media: International", "Gender", "Social Events"),
            dv.labels = c("Estimates"), string.se = "SE", string.stat = "t",
            show.se = TRUE, show.stat = TRUE, show.ci = FALSE, 
            show.re.var = FALSE, show.aic = FALSE,
            show.r2 = FALSE, show.icc = FALSE, show.obs = TRUE,
            CSS = css_theme("regression"), file = str_c(dir_graphs, "/table2.html"))
  
  # Papaja
  apa_lm <- papaja::apa_print(final_model) 
  apa_lm$table$predictor <- c("Intercept", "Alexithymia", "Anxiety", "Interpersonal (IRQ)", 
                              "Extraversion", "IUS", "Liberalism", "Age", "Covid Knowledge", 
                              "Covid Behavior", "Covid Media: Facebook", "Covid Media: Twitter", 
                              "Covid Media: TV", "Covid Media: USA", "Covid Media: International", "Gender", "Social Events")
  papaja::apa_table(apa_lm$table, 
                    caption = "Cross-validated regression model")
}
```

## Random forest

```{r random_forest}
## RF specific 
df_train_rf <- df_train_raw
df_test_rf <- df_test_raw

numCores <- parallel::detectCores() - 1 
cl <- parallel::makeCluster(numCores, type = "SOCK")
doSNOW::registerDoSNOW(cl) 

train_control <- trainControl(method = "cv", number = 10)
forest_fit <- train(covid_emo_distress_s ~ ., 
                data = df_train_rf, 
                method = "ranger", 
                trControl = train_control, 
                importance = "impurity")
parallel::stopCluster(cl) 

## Test correlation 
df_test_rf$model_fits <- predict(forest_fit, newdata = df_test_rf)
df_test_results_rf <- postResample(pred = df_test_rf$model_fits, obs = df_test_rf$covid_emo_distress_s)
cor_rf <- cor(df_test_rf$model_fits, df_test_rf$covid_emo_distress_s)
```

## Lasso

```{r lasso}
## Lasso specific
df_train_ls <- df_train_raw
df_test_ls <- df_test_raw

numCores <- parallel::detectCores() - 1 
cl <- parallel::makeCluster(numCores, type = "SOCK")
doSNOW::registerDoSNOW(cl) 

train_control <- trainControl(method = "cv", number = 10)
tune_control <- expand.grid(alpha = 1, lambda = 10^seq(-3, 3, length = 100))
lasso_fit <- train(covid_emo_distress_s ~ ., 
                data = df_train_ls, 
                method = "glmnet", 
                trControl = train_control,
                tuneGrid = tune_control)
parallel::stopCluster(cl)

## Test correlation 
df_test_ls$model_fits <- predict(lasso_fit, newdata = df_test_ls)
df_test_results_ls <- postResample(pred = df_test_ls$model_fits, obs = df_test_ls$covid_emo_distress_s)
cor_ls <- cor(df_test_ls$model_fits, df_test_ls$covid_emo_distress_s)
```

# Simple correlations

```{r simple_cor}
# Final model is cv
df_cv <- tidy(final_model, conf.int = TRUE) %>%
  mutate(p.labels = case_when(p.value <= .001 ~ "***", 
                              p.value <= .01 ~ "**", 
                              p.value <= .05 ~ "*", 
                              p.value > .05 ~ "n.s."))

# Simple relationships
df_simples <- df_models %>%
  pivot_longer(cols = -covid_emo_distress_s, names_to = "term", values_to = "value") %>% 
  group_by(term) %>% 
  do(tidy(cor.test(.$covid_emo_distress_s, .$value))) %>%
  select(term:conf.high) %>%
  mutate(p.labels = case_when(p.value <= .001 ~ "***", 
                              p.value <= .01 ~ "**", 
                              p.value <= .05 ~ "*", 
                              p.value > .05 ~ "n.s."))

# Full
df_cv_full <- df_cv %>%
  filter(term != "(Intercept)") %>%
  full_join(., df_simples, by = "term", suffix = c("_cross_valid", "_simple")) %>%
  mutate(grouping = case_when(term == "gad_total_s" ~ "mental_health", 
                              term == "ius_total_s" ~ "personality", 
                              term == "covid_media_usa_s" ~ "media", 
                              term == "covid_media_facebook_s" ~ "media", 
                              term == "covid_media_int_s" ~ "media", 
                              term == "covid_media_twitter_s" ~ "media", 
                              term == "covid_media_tv_s" ~ "media", 
                              term == "tas_total_s" ~ "mental_health", 
                              term == "irq_total_s" ~ "emotion_regulation", 
                              term == "covid_behave1_s" ~ "covid_measure", 
                              term == "covid_knowledge_s" ~ "covid_measure", 
                              term == "bfi_extraversion_s" ~ "personality", 
                              term == "gender" ~ "demographic", 
                              term == "age_s" ~ "demographic", 
                              term == "liberalism_s" ~ "demographic", 
                              term == "delta_social_s" ~ "social"), 
         error = case_when(((estimate_cross_valid < estimate_simple) & estimate_cross_valid > 0) ~ "over_estimate", 
                           ((estimate_cross_valid > estimate_simple) & estimate_cross_valid < 0) ~ "over_estimate", 
                           ((estimate_cross_valid < estimate_simple) & estimate_cross_valid < 0) ~ "under_estimate", 
                           ((estimate_cross_valid > estimate_simple) & estimate_cross_valid > 0) ~ "under_estimate"))

# Comparison
df_compare <- df_cv_full %>% filter(!is.na(grouping))

model_compare <- ggplot(df_compare, aes(x = estimate_cross_valid, y = estimate_simple)) + 
  geom_segment(aes(x = estimate_cross_valid, y = estimate_simple, xend = estimate_cross_valid, yend = estimate_cross_valid, linetype = error)) + 
  geom_point(aes(color = grouping)) + 
  #ggrepel::geom_label_repel(aes(label = term), size = 2) + 
  geom_abline(slope = 1, intercept = 0, linetype = 2) + 
  scale_x_continuous(name = "CV Estimates", limits = c(-.2, 0.7), breaks = seq(-.2, .7, .1)) + 
  scale_y_continuous(name = "Simple Estimates", limits = c(-.2, 0.7), breaks = seq(-.2, .7, .1)) + 
  coord_fixed() + 
  theme_classic() + 
  theme(text = element_text(size = 12))
model_compare
ggsave(filename = str_c(dir_graphs, "/fig1b.pdf"), plot = model_compare, width = 5, height = 4)

if (cor_output == TRUE) {
  ## Correlation plot
  final_variables <- as.data.frame(final_summary$coefficients) %>%
    tibble::rownames_to_column(var = "coefficient") %>%
    filter(coefficient != "(Intercept)") %>%
    pull(coefficient)
  
  cor_data <- cor(df_models %>% select(all_of(final_variables)), use = "na.or.complete")         # correlation of all variables 
  res1 <- cor.mtest(df_models %>% select(all_of(final_variables)), conf.level = .95)             # significance.
  
  ## Correlation network
  cor_df <- df_models %>% 
    select(all_of(final_variables), covid_emo_distress_s) %>% 
    correlate() %>%
    stretch() %>%
    mutate(x = case_when(x == "gad_total_s" ~ "Anxiety", 
                         x == "ius_total_s" ~ "IUS", 
                         x == "covid_media_usa_s" ~ "Media: USA", 
                         x == "covid_media_facebook_s" ~ "Media: Facebook", 
                         x == "covid_media_int_s" ~ "Media: International", 
                         x == "covid_media_twitter_s" ~ "Media: Twitter", 
                         x == "covid_media_tv_s" ~ "Media: TV", 
                         x == "tas_total_s" ~ "Alexithymia", 
                         x == "irq_total_s" ~ "Interpersonal (IRQ)", 
                         x == "covid_behave1_s" ~ "Covid Behavior", 
                         x == "covid_knowledge_s" ~ "Covid Knowledge", 
                         x == "bfi_extraversion_s" ~ "Extraversion", 
                         x == "gender" ~ "Gender", 
                         x == "age_s" ~ "Age", 
                         x == "liberalism_s" ~ "Liberalism", 
                         x == "social_events_s" ~ "Social Events", 
                         x == "covid_emo_distress_s" ~ "Covid Emotional Distress"), 
           y = case_when(y == "gad_total_s" ~ "Anxiety", 
                         y == "ius_total_s" ~ "IUS", 
                         y == "covid_media_usa_s" ~ "Media: USA", 
                         y == "covid_media_facebook_s" ~ "Media: Facebook", 
                         y == "covid_media_int_s" ~ "Media: International", 
                         y == "covid_media_twitter_s" ~ "Media: Twitter", 
                         y == "covid_media_tv_s" ~ "Media: TV", 
                         y == "tas_total_s" ~ "Alexithymia", 
                         y == "irq_total_s" ~ "Interpersonal (IRQ)", 
                         y == "covid_behave1_s" ~ "Covid Behavior", 
                         y == "covid_knowledge_s" ~ "Covid Knowledge", 
                         y == "bfi_extraversion_s" ~ "Extraversion", 
                         y == "gender" ~ "Gender", 
                         y == "age_s" ~ "Age", 
                         y == "liberalism_s" ~ "Liberalism", 
                         y == "social_events_s" ~ "Social Events", 
                         y == "covid_emo_distress_s" ~ "Covid Emotional Distress"))
  
  fig3_data <- cor_df %>%
    filter(abs(r) > .2) %>%
    graph_from_data_frame(directed = FALSE)
  
  fig3_plot <- ggraph(fig3_data) +
    geom_edge_link(aes(color = r)) +
    guides(edge_alpha = "none", edge_width = "none") +
    scale_edge_colour_gradientn(limits = c(-1, 1), colors = c("firebrick2", "dodgerblue2")) +
    geom_node_point(size = 3) +
    geom_node_text(aes(label = name), repel = TRUE) +
    theme_graph(base_family = 'Helvetica')
  fig3_plot
    
  ggsave(filename = str_c(dir_graphs, "/fig3.pdf"), plot = fig3_plot, width = 6, height = 4)
  ggsave(filename = str_c(dir_graphs, "/fig3.png"), plot = fig3_plot, width = 6.666, height = 5)
}
```
  
# Covid Emotion Distress Reliabilty

Code below investigates reliability of the covid emotion distress scale but the results are not printed to the Markdown. 

```{r reliability}
# Alpha
alpha_emo <- psych::alpha(df_emo_distress, keys = c("covid_emo_distress_1_2", "covid_emo_distress_2_6")) # keys are reverse coded items
#alpha_emo$total$raw_alpha

# Omega (ERF)
emo_r_matrix <- cor(df_emo_distress)
omega_emo <- omega(emo_r_matrix)  # bifactor factor solution 
#omegaSem(m = emo_r_matrix, n.obs = 948)

## Plot
#pdf(file = str_c(dir_graphs, "omega_plot.pdf"))
#omega(emo_r_matrix)
#dev.off()
```

# Session Info

```{r session_info}
sessionInfo()
```

# Appendix
Code used for all figures and analyses. 

```{r ref.label=knitr::all_labels(), echo = T, eval = F}
```